---
title: "Assignment_Four Geochemistry"
author: "Chiayu Tu"
date: "2/25/2022"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# Assignment 4
# clear the environment
rm(list = ls(all = TRUE))

# Set up my work space
setwd("C:/Users/tulou/Desktop/PSU_Classes/G_324/Assignment_Four")
# Run these libraries
library(tidyverse)
library(googlesheets4)
library(ggtern)
library(png)
library(ggplot2)

# authenticates your google account
gs4_auth()

# Link to googlesheet
myURL <- "https://docs.google.com/spreadsheets/d/1wyD7IWT0rouZI_Gw8eh4rHIO70dWdIvnEEZ_LOWIxWU/edit#gid=531586939"
```

## Statement of Problem

The "average" concentrations of the mantle and CI chondrites are observed by normalization, and the element names are presented on the x-axis and the normalized values are presented on the y-axis. All values will be presented as tibble.

## Methods

Using ggplot make a graph, let x is element name and y is the normalized values, there are two graphs, one is the average concentrations of mantle, the other one is the average concentrations of CI chondrites.
I retrieved the data from Google sheet the professor provided. I created a new table by joining different sheets from Google sheet. Using the values from ICP data divided by the data from mantle and CI chondrites.Plotting the average values on the graphs and seperate this data by the Age of different Unit.

\newpage

## Results

In these graphs, we can see the average element data of mantles. Through the samples of different Ages, we can see that the sample ages of most Units are concentrated in 11 and 16, and have relatively similar element averages. It can be seen that the Units of Ages 11 and 16 have a relatively large correlation.
```{r Average Mantle, echo=FALSE}
#Elements Table
ElementTableOne <- read_sheet(myURL, range = "ICP_INAAdata!A:AB", .name_repair = "unique") %>% 
  select(SampleID, Ba, Ce, Dy, Eu, Hf, La, Lu, Nb, Nd, Pb, Pr, Rb, Sm, Sr, Ta, Tb, Th, U, Y, Yb, Zr)
#view(ElementTableOne)

ElementTable <- pivot_longer(ElementTableOne, -c('SampleID'), names_to = 'element', values_to = 'values')
#view(ElementTable)

#Samples Table, include SampleID, Unit, Age, Latitude, Longitude
SampleTable <- read_sheet(myURL, range = "Samples!A:F", .name_repair = "unique") %>% 
  select(SampleID, Unit, Age, Latitude, Longitude)
#view(SampleTable)

SamplesTable <- inner_join(SampleTable, ElementTable)
#view(SamplesTable)

#Mantle Elements Table
MantleTable <- read_sheet(myURL, range = "Constants!G2:H26", .name_repair = "unique")
#view(MantleTable)

METable <- full_join(SamplesTable, MantleTable)
#view(METable)

#Average of Mantle Elements 
AveMETable <- function(num){
  if(num <= 12){
    METable %>% 
    as_tibble() %>% 
    rowwise() %>% 
    mutate(AveMentle = values / val) %>% 
    mutate(Age = round(Age, 0)) %>% 
    filter(, Age < num)
  }else{
    METable %>% 
    as_tibble() %>% 
    rowwise() %>% 
    mutate(AveMentle = values / val) %>% 
    mutate(Age = round(Age, 0)) %>%
    filter(, Age > num)
  }
}

#plot my data on the graph
MantleGraph <- function(name){
  ggplot(name, mapping = aes(x = element, y = val, group = element)) +
  ylim(0, 1000) +
  geom_point(, mapping = aes(x = element, y = AveMentle, group = Unit, fill = Unit)) +
  geom_line(aes(x= element, y = AveMentle, group = Unit, color = Unit)) +
  theme_bw() + 
  theme_showarrows() +
  weight_percent() +
  facet_wrap(~Age) + 
  theme(legend.position = "bottom")
}
MantleGraph(AveMETable(12))
MantleGraph(AveMETable(13))
```
\newpage

In these graphs, we can see the average element data of CI Chondrite.
```{r Average CI, echo=FALSE}
#CI Chondrite Elements Table 
CITable <- read_sheet(myURL, range = "Constants!J2:K26", .name_repair = "unique")
view(CITable)

CIETable <- full_join(SamplesTable, CITable)
view(CIETable)

#Average of CI Chondrite Elements
AveCIETable <- function(num){
  if(num <= 12){
    CIETable %>% 
    as_tibble() %>% 
    rowwise() %>% 
    mutate(AveCI = values / val) %>% 
    mutate(Age = round(Age, 0)) %>% 
    filter(, Age < num)
  }else{
    CIETable %>% 
    as_tibble() %>% 
    rowwise() %>% 
    mutate(AveCI = values / val) %>% 
    mutate(Age = round(Age, 0)) %>%
    filter(, Age > num)
  }
}

#plot my data on the graph(CI Chondrite)
CIGraph <- function(name){ 
  ggplot(name, mapping = aes(element, val, group = element)) +
  ylim(0, 1000) +
  geom_point(, mapping = aes(x = element, y = AveCI)) +
  geom_line(, mapping = aes(x = element, y = AveCI, group = Unit, color = Unit)) +
  theme_bw() + 
  theme_showarrows() +
  weight_percent() +
  facet_wrap(~Age) + 
  theme(legend.position = "bottom")
}

CIGraph(AveCIETable(12))
CIGraph(AveCIETable(13))
```
\newpage

## Conclusion

I set the x-axis to elements and the y-axis to their average, create a new table by joining the different tables, and find the average of the elements in common. Through these graphs, we can see that the samples obtained in the same Unit will also have different densities of elements due to Age or other factors. We can also guess that samples from different Units may originate from the same place by using similar element values.

